---
title: "DATA 624 Homework 7"
author: "Mario Pena"
date: "April 10, 2022"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(caret)
library(DataExplorer)
library(RANN)
```

### Exercise 6.2

Developing a model to predict permeability (see Sect. 1.4) could save significant resources for a pharmaceutical company, while at the same time more rapidly identifying molecules that have a sufficient permeability to become a drug:

**(a) Start R and use these commands to load the data:**

```{r}
library(AppliedPredictiveModeling)
data(permeability)
```

**The matrix `fingerprints` contains the 1,107 binary molecular predictors for the 165 compounds, while `permeability` contains permeability response.**

```{r}
dim(fingerprints)
head(permeability)
```


**(b) The fingerprint predictors indicate the presence or absence of substructures of a molecule and are often sparse meaning that relatively few of the molecules contain each substructure. Filter out the predictors that have low frequencies using the `nearZeroVar` function from the `caret` package. How many predictors are left for modeling? ** 

```{r}
filter <- nearZeroVar(fingerprints)
high_freq_pred <- fingerprints[, -filter]
dim(high_freq_pred)
```

After applying the `nerZeroVar` function and filtering out the low frequency predictors, we are left with 388 out of the original 1,107.


**(c) Split the data into a training and a test set, pre-process the data, and tune a `PLS` model. How many latent variables are optimal and what is the corresponding resampled estimate of R2?**

We will split the data into 80% train and 20% test.

```{r}
set.seed(123)
rows <- createDataPartition(permeability, p=0.8, list=FALSE) 
Train_X <- high_freq_pred[rows,]
Train_Y <- permeability[rows,]
Test_X <- high_freq_pred[-rows,]
Test_Y <- permeability[-rows,]

```

```{r}
set.seed(100)
plsTune <- train(Train_X, Train_Y,
method = "pls",
tuneLength = 20,
trControl = trainControl(method = "cv", number = 10),
preProc = c("center", "scale"))
```

```{r}
plsTune
```

The optimal number of latent variables is 7 with an R2 of 0.55.

```{r}
plot(plsTune, main="PLS Model")
```

It is also observed on the graph above that the lowest RMSE is achieved with 7 variables.


**(d) Predict the response for the test set. What is the test set estimate of R2? ** 

```{r}
plspredict <- predict(plsTune, Test_X)

postResample(pred=plspredict, obs = Test_Y)
```

The predictions on the test set yield an R2 of 0.36, which is lower than the training set R2.


**(e) Try building other models discussed in this chapter. Do any have better predictive performance? **

We will try building a ridge regression and elastic net model, which use penalization to reduce RMSE.

```{r}
ridgeGrid <- data.frame(.lambda = seq(0, .1, length = 15))

enetGrid <- expand.grid(.lambda = c(0, 0.01, .1), .fraction = seq(.05, 1, length = 20))
```

```{r warning=FALSE}
set.seed(100)
ridgeRegFit <- train(Train_X, Train_Y,
method = "ridge",
## Fit the model over many penalty values
tuneGrid = ridgeGrid,
trControl = trainControl(method = "cv", number = 10),
## put the predictors on the same scale
preProc = c("center", "scale"))
```

```{r}
ridgeRegFit
```

The ridge regression used an optimal penalty of lambda 0.1 which yielded a RMSE of 12.10.

```{r warning=F}
set.seed(100)
enetTune <- train(Train_X, Train_Y,
method = "enet",
tuneGrid = enetGrid,
trControl = trainControl(method = "cv", number = 10),
preProc = c("center", "scale"))
```

```{r}
enetTune
```

The elastic net used an optimal penalty of lambda 0.1 and fraction 0.35, which yielded a RMSE of 11.11 and R2 of 0.57.

It seems that the elastic net model has superior predictive performance than any of the previous models discussed in this exercise.

**(f) Would you recommend any of your models to replace the permeability laboratory experiment?  **

```{r}
enetpredict <- predict(enetTune, Test_X)

postResample(pred=enetpredict, obs = Test_Y)
```

With an R2 of about 0.4 for the predictions on our test set from our best model, I don't feel confident that we could replace the laboratory experiments with any of these models.


### Exercise 6.3

A chemical manufacturing process for a pharmaceutical product was discussed in Sect. 1.4. In this problem, the objective is to understand the relationship between biological measurements of the raw materials (predictors), measurements of the manufacturing process (predictors), and the response of product yield. Biological predictors cannot be changed but can be used to assess the quality of the raw material before processing. On the other hand, manufacturing process predictors can be changed in the manufacturing process. Improving product yield by 1 % will boost revenue by approximately one hundred thousand dollars per batch:

**(a) Start R and use these commands to load the data:  **

```{r}
data("ChemicalManufacturingProcess")
```

**The matrix `processPredictors` contains the 57 predictors (12 describing the input biological material and 45 describing the process predictors) for the 176 manufacturing runs. `yield` contains the percent yield for each run.  **

```{r}
dim(ChemicalManufacturingProcess)
```



**(b) A small percentage of cells in the predictor set contain missing values. Use an imputation function to fill in these missing values (e.g., see Sect. 3.8).  **

On the missing values plot below, we can observe that roughly 28 variables have missing values that range from 0.57% to 8.52% of values missing.

```{r}
plot_missing(ChemicalManufacturingProcess)
```

I will use the `preProcess` function from section 3.8 to fill in the missing values in the data.

```{r}
preProcValues <- preProcess(ChemicalManufacturingProcess, method = c("knnImpute"))
```

```{r}
data_imp <- predict(preProcValues, ChemicalManufacturingProcess)
```

As we can see below, we were able to successfully impute the missing values.

```{r}
plot_missing(data_imp)
```


**(c) Split the data into a training and a test set, pre-process the data, and tune a model of your choice from this chapter. What is the optimal value of the performance metric?  **

Based on the previous exercise and seeing that the elastic net model had better predictive performance, I will use this model to train the data.

We will split the data into 80% train and 20% test.

```{r}
set.seed(123)
index <- createDataPartition(data_imp$Yield, p=0.8, list=FALSE) 
Train <- data_imp[index, ]
Test <- data_imp[-index, ]

```

```{r warning=F}
set.seed(100)
enetTune2 <- train(x = Train[, 2:58], y = Train$Yield,
method = "enet",
tuneGrid = enetGrid,
trControl = trainControl(method = "cv", number = 10),
preProc = c("center", "scale"))
```

```{r}
enetTune2
plot(enetTune2)
```

The optimal value for the performance metric is a penalty of lambda 0.01 and fraction 0.35, which yields a RMSE of 0.62 and R2 of 0.63.


**(d) Predict the response for the test set. What is the value of the performance metric and how does this compare with the resampled performance metric on the training set?  **


```{r}
enetpredict2 <- predict(enetTune2, Test[, 2:58])

postResample(pred=enetpredict2, obs = Test$Yield)
```

The predictions on the test set yield a RMSE of 0.689, which is better than the one on the training set and R2 of 0.536, which is slightly lower than the one from the training set. Overall it seems that the model has performed very well on the test set.



**(e) Which predictors are most important in the model you have trained? Do either the biological or process predictors dominate the list?**

In this model, the most important predictor at the top of the list is "ManufacturingProcess32". However, there is not a predictor type that dominates the list, we have both biological and process predictors similarly as important in the list.

```{r}
important <- varImp(enetTune2)
important
```


**(f) Explore the relationships between each of the top predictors and the response. How could this information be helpful in improving yield in future runs of the manufacturing process?**

```{r}
corr_data1 <- data_imp %>%
  select(Yield, ManufacturingProcess32, BiologicalMaterial06, BiologicalMaterial03, ManufacturingProcess13, ManufacturingProcess36, ManufacturingProcess31, BiologicalMaterial02, ManufacturingProcess17, ManufacturingProcess09, BiologicalMaterial12)

corr_data2 <- data_imp %>%
  select(Yield, ManufacturingProcess06, BiologicalMaterial11, ManufacturingProcess33, ManufacturingProcess29, BiologicalMaterial04, ManufacturingProcess11, BiologicalMaterial01, BiologicalMaterial08, BiologicalMaterial09, ManufacturingProcess30)

plot_correlation(corr_data1)
plot_correlation(corr_data2)
```

As we can see on the correlation plots above, some of the most important predictor variables have strong positive or negative relationships with the response variable. However, there are a few variables that do not seem to have strong correlations with the response and yet were identified as important for the model. Knowing which variables are positively or negatively correlated to the yield can help improve it by making the necessary adjustments in the manufacturing process to increase yield.


